Broom: Safe Region-Based Memory Management for Dataflow Programs

### 2015-10-16 paper/overview.tex
Edits.

@@

>>>>

\subsection{Memory Safety}
\label{sec:memory-safety}

Memory safety in \name is predominantly due to its region type system.
However, the type system does not ensure the integrity of a small
subset of pointers, called the \emph{unsafe set}, for which runtime
verification is needed to guarantee safety. This is more a design
decision rather than an unfortunate inevitability. Rest of this
subsection presents an overview of the region type system, discusses
the unsafe set, and briefly describes the runtime mechanisms used to
ensure safety of references in the unsafe set.

At a fundamental level, the region type system's approach to memory
safety involves approximating (a). \emph{liveness}: the set of memory
regions that are live at any given program location, and (b).
\emph{outlives} relationships between them, and using this information
to preempt potentially unsafe references from being created. These
include:
\begin{itemize}
\item References into dead/non-existing regions,
\item References from longer-living stack regions to shorter-living ones, and 
\item References that escape transferable regions.
\end{itemize}
Unsurprisingly, the type system's approximation of liveness and
outlives information for stack regions is precise. However, for
transferable regions, statically determining liveness is equivalent to
determining if a C pointer is null in presence of dynamic memory
allocation, which is undecidable in general. Therefore, the region
type system has to be necessarily conservative in approximating the
liveness of transferable regions in order to ensure full memory
safety. Maintaining a good conservative approximation requires the
type system to employ techniques such as linear types~\cite{wadlerlt}
or unique pointers~\cite{rust}, which impose restrictions that are
often too hard to program around~\footnote{Internet has many tutorials
on how to \emph{fight} the borrow checker in Rust.}. This is
particularly unacceptable in the context of \name, as dataflow
programs are often run offline, and the concern for performance and
developer productivity outweighs the concern for safety.

\name's region type system therefore overapproximates the liveness
information for transferable regions, i.e., it might assume that a
transferable region is live when, in fact, it might have already been
transferred or freed. This necessarily leads to cases where creation of
unsafe references, and their subsquent dereferencing can go unnoticed
by the type system. Considering that stack regions can contain
pointers into transferable regions, the worst case cost of
overapproximating liveness is therefore the integrity of \emph{every}
pointer on stack.

Fortunately, \name avoids this worst case scenario. It does so by
restricting the unsafe set to only include $\RgnZ$ objects. In
particular, pointers \emph{into} transferable regions are excluded
from this set (hence safe). \name achieves this by imposing simple
restrictions on how $\RgnZ$ objects can be used to operate on
transferable regions. The following describes the $\RgnZ$ class, and
how $\RgnZ$ objects should be used to make allocations into, and to
read from transferable regions.

@@

\subsection{An Example}
